SubsidyExplorer - Step 0: Data Wrangling

### 2020-03-20 scripts/01_data_wrangling.Rmd
almost all data minus vessel list back in

IUU Consolidated vessel list
> NEED

@@

# Put them back together
fao_production_dat_edit <- fao_production_dat %>%
  dplyr::filter(iso3 != "EU") %>%
  bind_rows(fao_production_dat_eu) %>%
  bind_rows(fao_production_dat_us)

@@

# Aggregate by ISCCAP groups and export for time series plot
fao_production_dat_edit_groups <- fao_production_dat_edit %>%
  group_by(iso3, year, isscaap_group) %>%
  summarize(value = sum(value, na.rm = T))

@@

## Estimated landed price

@@

>>>>

**Other**
IUU Consolidated vessel list
> NEED

Cap and Tier Data from the US Proposal
> Citation for this?

@@

# Put them back together
fao_production_dat_edit <- fao_production_dat %>%
  dplyr::filter(iso3 != "EU") %>%
  bind_rows(fao_production_dat_eu) %>%
  bind_rows(fao_production_dat_us) %>%
  mutate(variable = "capture_production",
         units = "tonnes",
         source = "FAO (2019)")

@@

# Aggregate by ISCCAP groups and export for time series plot
fao_production_dat_edit_groups <- fao_production_dat_edit %>%
  group_by(iso3, year, isscaap_group, variable, units, source) %>%
  summarize(value = sum(value, na.rm = T))

@@

## Ex-vessel prices from Melnychuk et al. (2017)
```{r}
# Price data from Melynchuk et al. (2017)
melnychuk_price_dat <- read_csv(melnychuk_price_path) %>%
  clean_names(case = "snake") %>%
  dplyr::select(-entry) %>%
  dplyr::filter(year >= 2000 & year <= 2012) %>%
  group_by(year, scientific_name, asfis_species, isscaap_group, group_for_pairing) %>%
  summarize(exvessel = mean(exvessel, na.rm = T)) %>%
  ungroup()

# Pull data from 2012 to use for extrapolation to 2017
melnychuk_price_dat_mry <- melnychuk_price_dat %>%
  dplyr::filter(year == 2012) %>%
  rename(exvessel_for_extrap = exvessel) %>%
  dplyr::select(-year)

# CPI for 2012 - 2018 to convert USD
cpi_2012 <- 229.594
cpi_2013 <- 232.957
cpi_2014 <- 236.736
cpi_2015 <- 237.017
cpi_2016 <- 240.007
cpi_2017 <- 245.120

# Do extrapolation for missing years
melnychuk_price_dat_extrapolation <- melnychuk_price_dat %>%
  complete(nesting(scientific_name, asfis_species, isscaap_group, group_for_pairing), year = 2000:2017) %>%
  left_join(melnychuk_price_dat_mry, by = c("scientific_name", "asfis_species", "isscaap_group", "group_for_pairing")) %>%
  mutate(exvessel_adjusted = case_when(year <= 2012 ~ exvessel,
                                       year == 2013 ~ exvessel_for_extrap * (cpi_2013/cpi_2012),
                                       year == 2014 ~ exvessel_for_extrap * (cpi_2014/cpi_2012),
                                       year == 2015 ~ exvessel_for_extrap * (cpi_2015/cpi_2012),
                                       year == 2016 ~ exvessel_for_extrap * (cpi_2016/cpi_2012),
                                       year == 2017 ~ exvessel_for_extrap * (cpi_2017/cpi_2012))) %>%
  dplyr::select(-exvessel, -exvessel_for_extrap) %>%
  rename(exvessel = exvessel_adjusted)

# Get rid of NAs and save
melnychuk_price_dat_extrapolation <- melnychuk_price_dat_extrapolation %>%
  dplyr::filter(!is.nan(exvessel)) %>%
  mutate(variable = "ex_vessel_price",
         units = paste(year, "US$"),
         source = "Melnychuk et al. (2017)")

write_csv(melnychuk_price_dat_extrapolation, paste0(data_dir, "melnychuk_2017_price_tidy.csv"))

# Create averages by pairing group for each year
price_dat_group_year <- melnychuk_price_dat_extrapolation %>%
  group_by(year, group_for_pairing) %>%
  summarize(exvessel_group_avg = mean(exvessel, na.rm = T)) %>%
  ungroup()

# Create averages by pairing group (no year) 
price_dat_group <- melnychuk_price_dat_extrapolation %>%
  group_by(group_for_pairing) %>%
  summarize(exvessel_group_avg_allyears = mean(exvessel, na.rm = T)) %>%
  ungroup()
```

## Estimated landed value from FAO/Melnychuk et al. (2017)

```{r}
# Combine landings and price data
combine_catch_price_dat <- fao_production_dat_edit %>%
  left_join(melnychuk_price_dat_extrapolation %>% dplyr::select(-isscaap_group, -variable, -source, -units), by = c("year", "scientific_name", "asfis_species", "isscaap_group" = "group_for_pairing")) %>%
  left_join(price_dat_group_year, by = c("year", "isscaap_group" = "group_for_pairing")) %>%
  left_join(price_dat_group, by = c("isscaap_group" = "group_for_pairing")) %>%
  mutate(exvessel_inclusive = case_when(!is.na(exvessel) ~ exvessel,
                                        !is.na(exvessel_group_avg) ~ exvessel_group_avg,
                                        !is.na(exvessel_group_avg_allyears) ~ exvessel_group_avg_allyears))

# Look at NAs - 8498 entries
combine_nas <- combine_catch_price_dat %>%
  dplyr::filter(is.na(exvessel_inclusive))

# Set price to 0 for those
combine_catch_price_dat$exvessel_inclusive[is.na(combine_catch_price_dat$exvessel_inclusive)] <- 0

# Calculate revenues and aggregate by ISSCAAP group
landed_value_dat <- combine_catch_price_dat %>%
  dplyr::select(-exvessel, -exvessel_group_avg, -exvessel_group_avg_allyears) %>%
  rename(exvessel = exvessel_inclusive) %>%
  mutate(landed_value = value*exvessel) %>%
  group_by(iso3, year, isscaap_group) %>%
  summarize(value = sum(landed_value, na.rm = T)) %>%
  mutate(variable = "landed_value",
         units = paste(year, "US$"),
         source = "FAO (2019); Melnychuk et al. (2017)")

write_csv(landed_value_dat, paste0(data_dir, "estimated_landed_value_isscaap_groups_tidy.csv"))
write_csv(landed_value_dat, paste0(app_data_dir, "estimated_landed_value_isscaap_groups_tidy.csv"))
```

@@

# Section 4 - Other Data

## Cap and Tier Data from the US Proposal
```{r}
cap_tier_dat <- read_csv(cap_tier_path) %>%
  mutate(iso3 = case_when(country == "Eswatini" ~ "SWZ",
                          country == "EU" ~ "EU",
                          TRUE ~ countrycode(country, "country.name", "iso3c")))
```
